---
title: Classical nights
author: Thomas Klebel
date: '2017-11-04'
slug: classical-nights
draft: true
categories:
  - Coding
tags:
  - R
---

One of the many inspirations to start blogging was a post by Maëlle Salmon about
the [program of Radio Swiss classic](http://www.masalmon.eu/2017/04/23/radioswissclassic/). In her post
Maëlle scraped data about names of composers, durations of pieces, and titles
of pieces, from which she tries to infer the tempo of pieces. From this data
what I found most interesting was her list about which composers appear most
frequently. This list supports the notion of a conservative canon of what counts
to "classical music".

My general impression about the selection of pieces for performance within the
classical music world is that pieces are selected quite conservatively. There
is basically a canon of composers from which pieces are repeatedly performed.
This canon seems to be quite stable, ranging from J.S.Bach and A.Vivaldi,
representing what counts as "early music", continuing with the First Viennes
School (Haydn, Mozart, Beethoven), through composers of the romantic phase (like 
Schubert, Schumann, Brahms, Verdi, Dvořák, Liszt, Tchaikovsky, Bruckner, Wagner),
followed by more "modern" composers such as Richard Strauss and Mahler, together
with Debussy, Chopin and Ravel. Obviously we could list a few more composers,
but my argument is that what counts to classical music more often than not starts
with exemplary representatives of "early music" (Bach and Vivaldi) and ends
with the beginning of the Second Viennese School (Schönberg, Webern) and their
departure from tonal music.

There would be many ways for trying to proof this hypothesis:

- taking a look at the literature, whether the topic of a "conservative canon" is
a thing there
- looking at radio stations
- looking at performances of typical orchestras or at typical venues

and probably many more.

For startes I wanted to compare what is being aired on Radio Swiss classic
with the program of an Austrian counterpart - [radio klassik Stephansdom](https://radioklassik.at). Since the internal layout of their
page is a lot more complicated than the one of Radio Swiss classic^[
The program is organised into different shows throughout the day, and not a
continuous stream of music], I will for starters just download pieces from during
night time (the programme is called "Nacht", german for "night"), which consists
of 5-7 hours of uninterrupted music. This should compare nicely to the part
of Maëlle's data on pieces during night time.

# Scraping the data

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(rvest)


base_url <- "https://radioklassik.at/programm/sendungsdetails/?sdk=nacht&sdd="

# generate dates
seq(ymd("2017-01-01"), ymd("2017-02-01"), 1) %>% 
  str_replace_all("-", ".")

# second url
# https://radioklassik.at/programm/sendungsdetails/?sdk=radio_klassik_stephansdom_nacht&sdd=14.10.2016
# 
# https://radioklassik.at/programm/sendungsdetails/?sdk=notturno&sdd=17.08.2013
# 
# https://radioklassik.at/programm/sendungsdetails/?sdk=nacht&sdd=14.01.2017

getwd()

```

```{r, eval=FALSE, echo=FALSE}
library("rvest")
library("purrr")
library("stringi")
library("lubridate")
library("tidyverse")

# using purrr::safely is cool because if the page 
# is broken you get NULL as an output
s_read_html <- purrr::safely(read_html)

# helper for brevity
xtract_nodes <- function(node, css) {
  html_nodes(node, css) %>% html_text(trim = TRUE)
}

get_one_day_program <- function(date=Sys.Date(),
                                base_url="http://www.radioswissclassic.ch/en/music-programme/search/%s",
                                pb=NULL) {
  # progress bar magic!
  if (!is.null(pb)) pb$tick()$print()

  # that's the part where you're nice towards the website
  Sys.sleep(sample(5,1)) # ideally, make this sample(5,1)

  date <- ymd(date) # handles case where input is character ISO date

  pg <- s_read_html(sprintf(base_url, format(date, "%Y%m%d")))

  if (!is.null(pg$result)) {
# go read Bob Rudis' post, in particular to read about
# the extra selector for "playlist"
    dplyr::data_frame(

      date = date,
      time = xtract_nodes(pg$result, 'div[class="playlist"] *
                                            span[class="time hidden-xs"]') %>% hm(),
      datetime = update(date, hour = hour(time), minute = minute(time)),
      artist = xtract_nodes(pg$result, 'div[class="playlist"] * span[class="titletag"]'),
      title = xtract_nodes(pg$result, 'div[class="playlist"] * span[class="artist"]')

    )

  } else {
    closeAllConnections()
    NULL
  }

}

search_dates <- seq(from = ymd("2008-09-01"), to = ymd("2017-04-22"), by = "1 day")

# how you can make the progress bar work,
# estimate the time necessary for 5 requests
pb <- dplyr::progress_estimated(length(search_dates))
programs_df <- map_df(search_dates, get_one_day_program, pb = pb)
programs_df <- programs_df %>%
   dplyr::select(-time) %>%
   dplyr::mutate(datetime = force_tz(datetime, tz = "Europe/Vienna"))


write_rds(programs_df, "2017-11_swiss_radio.rds", compress = "gz")
```



```{r}
programs_df <- read_rds("2017-11_swiss_radio.rds")
programs_df %>% 
  filter(artist == "Antonin Dvorak") %>% 
  count(title) %>% 
  arrange(desc(n)) %>% 
  head(30) %>% 
  mutate(index = 1:n()) %>% 
  select(index, everything()) %>% 
  knitr::kable()
```


```{r}
programs_df %>% 
  mutate(opera = str_detect(title, "Opera"),
         symphony = str_detect(title, "Symphony"),
         concerto = str_detect(title, "Concerto"),
         time = hour(datetime)) %>% 
  group_by(time) %>% 
  summarise(n = n(),
            operas = sum(opera)/n,
            symphonies = sum(symphony)/n,
            concertos = sum(concerto)/n) %>% 
  select(-n) %>% 
  gather(type, val, -time) %>% 
  ggplot(aes(time, val, colour = type)) +
  geom_line() +
  hrbrmisc::theme_hrbrmstr()

# bad approach, should normalise over duration
```


```{r}
pdata <- programs_df %>% 
  filter(artist %in% c("Johannes Brahms", "Wolfgang Amadeus Mozart")) %>% 
  mutate(opera = str_detect(title, "Opera"),
         symphony = str_detect(title, "Symphony"),
         concerto = str_detect(title, "Concerto"),
         time = hour(datetime)) %>% 
  group_by(artist) %>% 
  summarise(n = n(),
            operas = sum(opera)/n,
            symphonies = sum(symphony)/n,
            concertos = sum(concerto)/n) %>% 
  select(-n) %>% 
  gather(type, val, -artist)

pdata %>% 
  ggplot(aes(type, val, fill = artist)) +
  geom_col(position = "dodge") +
  coord_flip() +
  hrbrmisc::theme_hrbrmstr()
```


