---
title: Scraping Composers from Wikipedia
author: Thomas Klebel
date: '2017-10-30'
slug: scraping-composers-from-wikipedia
draft: false
categories:
  - Coding
tags:
  - R
  - Classical Music
---

In this post I will demonstrate how to download data about classical composers 
from wikipedia. Admittedly, I'm not an expert in web-scraping. The approach I
will demonstrate works for me, but there might be neater or more robust ways
to acheive the same result.

Wikipedia hast quite extensive lists on composerse of classical music. For the
purpose of informing my last post on the music played on radio swiss classic,
I wanted to add information on when the highest ranking composers lived.

I will section the post according to wikipedia's structure into baroque, 
classical era, romantic era and 20th century.

As for packages, we will use the `tidyverse` for general functions and `rvest`
for scraping the pages.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(rvest)

# convenience function to print first few lines of data.frames
print_composers <- function(df, n = 10) df %>% head(n) %>% knitr::kable()
```


# Baroque period (1600-1760)
For this period, the data is in a simple list of the following format:
`'First name' 'Surname' ('date of birth' - 'date of death')`. 

```{r}
link_baroque <- "https://en.wikipedia.org/wiki/List_of_Baroque_composers"
pg <- read_html(link_baroque)

baroque_composers <- pg %>% 
  html_nodes("ul") %>%  
  .[2:9] %>%  # only those lists are of interet to us
  html_nodes("li") %>% 
  html_text() %>% 
  data_frame(complete = .)
```

In the next step we will extract the name, date of birth and of death. For this
I will define a function, since we can re-use the code for other eras too. In
this function we leverage `str_extract` from the `stringr`-package and some
regex-expressions, which I verbalise with inline comments.

```{r}
extract_info <- function(df) {
  df %>% 
    mutate(
      # regex: start from beginning until there is a " (", which is omitted
      composer = str_extract(complete, "^.*?(?=\\s\\()"), 
      # regex: extract the first four digits
      year_of_birth = str_extract(complete, "\\d{4}"),
      # regex: extract the last four digits, either before ")" or before "?)"
      year_of_death = str_extract(complete, "\\d{4}(?=\\??\\))")) %>% 
    select(-complete) %>% 
    mutate_at(c("year_of_birth", "year_of_death"), as.numeric)
}
```


After defining the function we can use it on our baroque composers: 

```{r}
baroque_composers <- baroque_composers %>% 
  extract_info()

print_composers(baroque_composers)
```

There are a few edge cases, which I didn't catch, but this should be sufficient
four our purposes.


# Classical era
For the classical era we can use the same procedure as above.

```{r}
link_classical <- "https://en.wikipedia.org/wiki/List_of_Classical-era_composers"
pg <- read_html(link_classical)

classical_composers <- pg %>% 
  html_nodes("ul") %>%  
  .[2:6] %>%  # only those lists are of interet to us
  html_nodes("li") %>% 
  html_text() %>% 
  data_frame(complete = .) %>% 
  extract_info()

print_composers(classical_composers)
```


# Romantic era
For the composers of the romantic era we need a different approach, since they
are stored in a html-table. Luckily the tables can be easily accessed via the
css-class `table.wikitable`. We convert the content to a `data_frame` by mapping
`html_table` over all five tables. In the next step I fix a problem with the
date of birth of Scott Joplin.

```{r}
link_romantic <- "https://en.wikipedia.org/wiki/List_of_Romantic-era_composers"
pg <- read_html(link_romantic)

composers_romantic <- pg %>% 
  html_nodes("table.wikitable") %>% 
  map(html_table) %>% 
  map_at(5, ~mutate_at(.x,
                       c("Date born"),
                       function(x) {str_extract(x, "\\d{4}") %>% parse_number()})) %>%
  bind_rows() %>% 
  rename(composer = Name,
         year_of_birth = "Date born", 
         year_of_death = "Date died")
```

# 20th century
The table for composers of the 20th century is slightly more elaborate, which 
complicates things for use. In order to be able to sort by surname, the table
contains a sortkey ("Beethoven, Ludwig van" in case of Beethoven). When 
converting to a data_frame, the name of the composer ends up being 
"Beethoven, Ludwig vanLudwig van Beethoven". 

```{r}
link_composers_20th <- "https://en.wikipedia.org/wiki/List_of_20th-century_classical_composers"
pg <- read_html(link_composers_20th)
  
table_20 <- pg %>% 
  html_node("table.wikitable") %>%
  html_table(fill = TRUE)

print_composers(table_20, 3)
```

I found no better way than to accept this, extract the sortkeys separately, and
then delete the sortkeys from the column with
`str_replace(composer, sortkeys, "")`. If there is a better way, I would be glad
if someone could point it out in the comments.

```{r}

sortkeys <- pg %>% 
  html_node("table.wikitable") %>%
  html_nodes(css = 'span[class="sortkey"]') %>% 
  html_text()

composers_20th <- table_20 %>% 
  select(composer = Name, contains("Year")) %>% 
  rename(year_of_birth = "Year of birth", 
         year_of_death = "Year of death") %>% 
  mutate(sortkeys = sortkeys,
         composer = str_replace(composer, sortkeys, ""),
         year_of_birth = parse_number(year_of_birth)) %>% 
  select(-sortkeys)

print_composers(composers_20th)
```


# Combine all composers
```{r}
all_composers <- bind_rows(
  baroque_composers,
  classical_composers,
  composers_romantic,
  composers_20th
) 
```

Until now this post consits only of code and tables, so lets use our newly
gathered data and create a nice graph. Since some composers are listed for two
periods, because they lived during their transition, we should remove these
"duplicates".

```{r}
all_composers %>% 
  distinct(composer, year_of_birth, year_of_death) %>%
  ggplot(aes(year_of_birth)) +
  geom_freqpoly(bins = 40) +
  hrbrmisc::theme_hrbrmstr() +
  labs(x = "Year of birth", y = "# of composers",
       title = "Birth-years for classical composers\nlisted on wikipedia")
```



# Still to do
- Add dates for periods, some explanation which composers are associated with
borders of periods.
- Think and write about interpretation of graph.

